\section{Your Task}\label{sec:task}

\noindent You will be writing a program called \texttt{identify} that
identifies the top $k$ likely authors that authored an anonymous sample of text
passed into \texttt{stdin}. Your program will need to handle the following
command-line options:

\begin{itemize}
  \item \texttt{-d} : Specify path to database of authors and texts. The
    default is \texttt{lib.db}.
  \item \texttt{-n} : Specify path to file of noise words to filter out. The
    default is \texttt{noise.txt}.
  \item \texttt{-k} : Specify the number of matches. The default is 5.
  \item \texttt{-l} : Specify the number of noise words to filter out. The
    default is 100.
  \item \texttt{-e} : Set the distance metric to use as the Euclidean distance.
    This should be the default metric if no other distance metric is specified.
  \item \texttt{-m} : Set the distance metric to use as the Manhattan distance.
  \item \texttt{-c} : Set the distance metric to use as the cosine distance.
  \item \texttt{-h} : Display program help and usage.
\end{itemize}

You will be provided with a file, \texttt{metric.h}, that enumerates each of the
possible distance metrics and contains an array of strings, one for each
distance metric. Your main program, implemented in \texttt{identify.c}, should
do the following:

\begin{itemize}
  \item Create a new \texttt{Text} that contains words in the noise word file.
    This will be the noise word \texttt{Text} that contains words to filter out.
    Make sure that each word in the noise word \texttt{Text} is lowercased.
  \item Create a new anonymous \texttt{Text} from text passed in to
    \texttt{stdin}. This is the text that you want to identify. Remember to
    filter out all the noise words in the noise word \texttt{Text} that you just
    created.
  \item Open up the database of authors and text files. This file, by default,
    should be \texttt{lib.db}. The first line of this file indicates the number
    of author/text pairs. We'll refer to this number as $n$.
  \item Create a priority queue that can hold up to $n$ elements. This will be
    used to efficiently find the top $k$ likely authors who authored the
    anonymous text.
  \item Process the rest of the lines in the database file in pairs. You will
    need to do the following:
    \begin{itemize}
      \item Use \texttt{fgets()} to read in the name of an author, then use
        \texttt{fgets()} again to read in the path to the file of text that
        author authored. You will need to remove the newline character that
        \texttt{fgets()} preserves.
      \item Open up the author's file of text and use it to create a new
        \texttt{Text} instance. If the file could not be opened, continue to the
        next line of the database file. Remember to turn each parsed word to
        lowercase before checking if the word should be added to the
        \texttt{Text}.
      \item Compute the distance between the author's \texttt{Text} and the
        anonymous \texttt{Text}, filtering out words in the noise word
        \texttt{Text}.
      \item Enqueue the author's name and the computed distance into the
        priority queue.
    \end{itemize}
  \item Get the top $k$ likely matches by dequeueing the priority queue. Note
    that there may be less $k$ entries in the priority queue, depending on
    whether or not a file could be processed.
\end{itemize}

Example program output is provided below. Note that the top $k$ matches your
program finds should match the reference program exactly, but the displayed
distance may differ slightly due to floating point rounding errors.

\begin{shlisting}{}
$ ./identify -k 10 < texts/william-shakespeare.txt
Top 10, metric: Euclidean distance, noise limit: 100
1) William Shakespeare [0.000000000000000]
2) William D. McClintock [0.026516404305561]
3) Dante Alighieri [0.026865145178097]
4) Edgar Allan Poe [0.027829100437714]
5) Johann Wolfgang von Goethe [0.027873601844574]
6) Henry Timrod [0.028370953256889]
7) Lew Wallace [0.028546232963706]
8) Saxo Grammaticus [0.029282174356059]
9) Rudyard Kipling [0.029385152618790]
10) Various [0.029410943073263]
\end{shlisting}
