\section{Shell Sort}

\epigraph{\emph{There are two ways of constructing a software design.
    One way is to make it so simple that there are obviously no
    deficiencies, and the other way is to make it so complicated that
    there are no obvious deficiencies. The first method is far more
difficult.}}{---C.A.R. Hoare}

\noindent Donald L. Shell (March 1, 1924--November 2, 2015) was an
American computer scientist who designed the Shell sort sorting
algorithm. He earned his Ph.D. in Mathematics from the University of
Cincinnati in 1959, and published the Shell Sort algorithm in the
Communications of the ACM in July that same year.

Shell Sort is a variation of Insertion Sort and first sorts pairs of
elements which are far apart from each other. The distance between these
pairs of elements is called a \emph{gap}. Each iteration of Shell Sort
decreases the gap until a gap of 1 is used. At this point, the elements
are sorted. What is the expected time complexity of Shell Sort? It
depends entirely upon the gap sequence.

Any gap sequence that ends with a gap of 1 will produce a sorted array.
That said, a gap sequences with either too many or too few gaps should
be avoided as they either slow down passes through the array, or incur
additional overhead. Perhaps one of the best known gap sequences is the
sequence proposed by Vaughan Pratt in 1971 that is $\operatorname{O}(n
\log^2 n)$. American computer scientist Donald Knuth later proposed his
own gap sequence based off Pratt's in 1973 that is
$\operatorname{O}(n^{\frac{3}{2}})$. Robert Sedgewick eventually
proposed his own gap sequence in 1982. His gap sequence is
$\operatorname{O}(n^{\frac{4}{3}})$.

The following is the pseudocode for Shell Sort using Knuth's gap
sequence. The sequence is produced using a \emph{generator} in
\textbf{Python}. You are expected to implement exactly this gap
sequence. Note that \textbf{C} \emph{does not} have generators, so you
will need to find a way around this issue. For each \texttt{gap} in the
gap sequence, the sort compares all the pairs in array \texttt{A} that
are \texttt{gap} indices away from each other, shifting back array
elements that are in the wrong order.

\begin{pylisting}{Shell Sort in Python}
from math import log

def gaps(n: int):
    for i in range(int(log(3 + 2 * n) / log(3)), 0, -1):
        yield (3**i - 1) // 2

def shell_sort(A: list):
    for gap in gaps(len(A)):
        for i in range(gap, len(A)):
            j = i
            temp = A[i]
            while j >= gap and temp < A[j - gap]:
                A[j] = A[j - gap]
                j -= gap
            A[j] = temp
\end{pylisting}

From the definition of the gap sequence generator, you should see that
gaps are computed as $\lfloor \frac{3^k - 1}{2} \rfloor$ and the largest
$k$ is $\lfloor \frac{\log(2n + 3)}{\log(3)} \rfloor$, where $n$ is the
length of the array to sort. You be iterating over the gap sequence
starting with the largest $k$ first, making your way down to a gap size
of 1.
