\section{Specifics}

\epigraph{\emph{Vielleicht sind alle Drachen unseres Lebens
  Prinzessinnen, die nur darauf warten uns einmal sch\"on und mutig zu
sehen. Vielleicht ist alles Schreckliche im Grunde das Hilflose, das von
uns Hilfe will.}}{---Rainer Maria Rilke}

\noindent The following subsections cover the requirements of your test harness.
It is crucial that you follow the instructions.

\subsection{Command-line Options}

Your test harness must support any combination of the following
command-line options:

\begin{itemize}
  \item \texttt{-a}\ : Enables \emph{all} sorting algorithms.
  \item \texttt{-b}\ : Enables Bubble Sort.
  \item \texttt{-s}\ : Enables Shell Sort.
  \item \texttt{-q}\ : Enables the Quicksort that utilizes a stack.
  \item \texttt{-Q}\ : Enables the Quicksort that utilizes a queue.
  \item \texttt{-r seed}\ : Set the random seed to \texttt{seed}.
    The \emph{default} seed should be 13371453.
  \item \texttt{-n size}\ : Set the array size to \texttt{size}. The
    \emph{default} size should be 100.
  \item \texttt{-p elements}\ : Print out \texttt{elements} number of
    elements from the array. The \emph{default} number of elements to
    print out should be 100. \textcolor{red}{If the size of the array is
      less than the specified number of elements to print, print out the
    entire array and nothing more.}
\end{itemize}

It is important to read this \emph{carefully}. None of these options are
\emph{exclusive} of any other (you may specify any number of them,
including \emph{zero}). The most natural data structure for this
problem is a \emph{set}.
