\section{\LaTeX{} (\texttt{pdflatex})}

What is \LaTeX? It is a tool used by scientists, mathematicians and
engineers to produce high quality documents. It is especially
useful when your documents have mathematics in them.

Here is an example minimal \LaTeX document. There is a much more
complete example in the course resources repository. If your GitLab repository
has been created, then you should have access to this repository. It contains
the materials for each assignment: header files, source examples, and example
programs. You can access the repository here:

\centerline{\url{https://git.ucsc.edu/cse13s/winter2022/resources}}

\begin{verbatim}
\documentclass[11pt]{article} % Document type
\title{Example Document}
\author{Conrad Veidt}
\date{\today} % Sets the date to \today, or any date you specify
\begin{document}\maketitle % Start the document
For the left Riemann sum, approximating the function by its value at the
left-end point gives multiple rectangles with base $\Delta x$ and height
$f(a + i\Delta x)$. Doing this for $i = 0, 1, \ldots , n-1$, and adding
up the resulting areas gives
$$
A_\text{left} =
\Delta x \left[f(a) + f(a + \Delta x)
+ f(a+ 2 \, \Delta x)+ \cdots+f(b - \Delta x)\right].
$$
\end{document}
\end{verbatim}

You will need to install \LaTeX{} before you can start using it. You will most
likely want the full installation:

\begin{shlisting}{}
$ sudo apt install texlive-full
\end{shlisting}

There is a complete example in the resources repository that you can modify.
\LaTeX{} is a compiled language, just like \C{}. That means if you make a syntax
error it will complain!

Do you have to use \LaTeX? No, but you do have to use \texttt{gnuplot}.
Life will be much easier if you take a few minutes to learn and use
\LaTeX{}. All you need to do is to mimic what you find in the example
documents.
