\documentclass{article}
\usepackage{fullpage,fourier}
\usepackage{color}
\usepackage[x11names]{xcolor}
\usepackage{xspace}
\usepackage{url}

\newcommand{\surl}[1]{{\fontsize{9}{11}\selectfont\url{#1}}}

\title{Acknowledgment of Academic Honesty Policy}
\author{Prof.\xspace Darrell Long \\
CSE 13S}
\date{Winter 2021}

\begin{document}\maketitle
I have read and understand the policy on academic dishonesty (more
commonly referred to as ``cheating'') as described in the syllabus
for the Winter 2021 offering of CSE 13S
and the University policy on academic honesty, listed at
\surl{https://ue.ucsc.edu/academic-misconduct.html}.  I have asked for
clarification of any details of the academic honesty policy that I
did not understand.

In particular, I understand that obtaining solutions, in whole or
in part (text or source code) to graded class assignments from any
source---including work done by former or current Computer Science
students, professors, tutors, other humans, animals, zombies, space
aliens, reptilians from the core of the Hollow Earth,
materials found on the Internet, \emph{et cetera}---that
is not explicitly sanctioned by Prof.\xspace Long or the teaching assistants is unacceptable.
I also understand that \emph{providing} full or partial solutions
to anyone else, whether in this class or not, is likewise unacceptable.
I understand that all work I submit must be exclusively my own,
with the exception of any team projects during which I am allowed
to collaborate only with my assigned partners.  Information and
assistance from the professor, teaching assistants, assigned project partners, course
textbooks, course web site, and code and documentation
(\emph{e.g.},~\texttt{man} pages) \emph{supplied with the \textbf{C} Standard
Library} (\texttt{libc})
may be freely used without acknowledgment.  I agree to document
\textbf{\emph{all other sources of aid}}, even if they are considered
acceptable, for each submitted assignment.

This policy is a clarification of university policy:
\begin{quote}
As defined in the Handbook, plagiarism includes not only the literal
incorporation---without appropriate citation---of phrases and
sentences from the work of another person, but also the use of
``views, opinions, or insights of another without acknowledgment''
and ``paraphrasing the characteristic or original phraseology,
metaphor, or other literary device of another without proper
attribution.''
\end{quote}


\textcolor{red}{There is no flexibility in this policy. If you
cheat, you fail, and your academic career is likely to be prematurely
terminated. The reason that you cheated is entirely immaterial.}
This policy applies equally to all parties, regardless of whether
they are transmitting or receiving the material in question.

\vspace{0.6in}

\noindent
\begin{tabular}{l@{\,}ll@{\,}l}
\vspace{0.5in} Print your name: & \rule{2.5in}{1pt} & CruzID: & \rule{1.2in}{1pt}\small{\texttt{@ucsc.edu}} \\
	 Sign your name:  & \rule{2.5in}{1pt} & Date:   & \rule{1.2in}{1pt}
\end{tabular}

\vspace{0.4in}
\noindent
\centerline{\fbox{
\begin{minipage}{6in}
A signed copy of this agreement \emph{must} be turned in \emph{to the professor} before
we will accept any assignment submissions from you.  This may be done before or after
class, or during office hours.
\end{minipage}
}}

\end{document}
