\section{Lexical Analysis with Regular Expressions}

\noindent You will need a function to parse out the words from samples of text,
which will be passed to you in the form of input streams. You are concerned with
normal words, \emph{contractions}, and \emph{hyphenations}. A normal word is any
sequence of one or more characters that are part of your regular expression word
character set. Your word character set should contain characters from a--z and
A--Z. You should also accept contractions like ``don't'' and ``y'all've'' and
hyphenations like ``pseudo-code'' and ``move-to-front''.

You will need to write your own \emph{regular expression} for a word, utilizing
the \texttt{regex.h} library to lexically analyze the input stream for words.
You will be given a parsing module that lexically analyzes the input stream
using your regular expression. You are not required to use the module itself,
but it is \emph{mandatory} that you parse through an input stream for words
using at least one regular expression. The interface for the parsing module will
be in \texttt{parser.h} and its implementation will be in \texttt{parser.c}.

The function \texttt{next\_word()} requires two inputs, the input stream
\texttt{infile}, and a pointer to a compiled regular expression,
\texttt{word\_regex}. Notice the word \emph{compiled}: you must first compile
your regular expression using \texttt{regcomp()} before passing it to the
function. Here is a small program that prints out words input to \texttt{stdin}
using the parsing module. In the program, the regular expression for a word
matches one or more lowercase and uppercase letters. The regular expression you
will have to write for your assignment will be more complex than the one
displayed here, as it is just an example.

\begin{clisting}{Example program using the parsing module.}
#include "parser.h"
#include <regex.h>
#include <stdio.h>

#define WORD "[a-zA-Z]+"

int main(void) {
    regex_t re;
    if (regcomp(&re, WORD, REG_EXTENDED)) {
        fprintf(stderr, "Failed to compile regex.\n");
        return 1;
    }

    char *word = NULL;
    while ((word = next_word(stdin, &re)) != NULL) {
        printf("Word: %s\n", word);
    }

    regfree(&re);
    return 0;
}
\end{clisting}
