\section{An SS Library}
\epigraph{\emph{If you think cryptography is the answer to your problem, then you don't know what your problem is.}}{---Peter G.\xspace Neumann}

\noindent
\begin{funcdoc}{void ss\_make\_pub(mpz\_t p, mpz\_t q, mpz\_t n, uint64\_t nbits,
  uint64\_t iters)}
  Creates parts of a new SS public key: two large primes \texttt{p} and
  \texttt{q}, and \texttt{n} computed as $\texttt{p} * \texttt{p} * \texttt{q}$.
  Begin by creating primes \texttt{p} and \texttt{q} using \texttt{make\_prime()}.
  We first need to decide the number of bits that go to each prime
  respectively such that $\log_2(\texttt{n}) \ge \texttt{nbits}$. Let the
  number of bits for \texttt{p} be a random number in the range $\left
  [\texttt{nbits}/{5}, (2 \times \texttt{nbits})/{5} \right )$. Recall that $n
  = p^2 \times q$: the bits from \texttt{p} will be contributed to \texttt{n}
  twice, the remaining bits will go to \texttt{q}. The number of Miller-Rabin
  iterations is specified by \texttt{iters}. You should obtain this random
  number using \texttt{random()} and check that $\texttt{p} \nmid \texttt{q} -
  1$ and $\texttt{q} \nmid \texttt{p} - 1$.
\end{funcdoc}

\begin{funcdoc}{void ss\_write\_pub(mpz\_t n, char username[], FILE *pbfile)}
  Writes a public SS key to \texttt{pbfile}. The format of a public key
  should be \texttt{n}, then the
  \texttt{username}, each of which are written with a trailing newline.
  The value \texttt{n} should be written as
  a \emph{hexstring}. See the GMP functions for
  \href{https://gmplib.org/manual/Formatted-Output-Functions}{formatted
  output} for help with writing hexstrings.
\end{funcdoc}

\begin{funcdoc}{void ss\_read\_pub(mpz\_t n,
char username[], FILE *pbfile)}
  Reads a public SS key from \texttt{pbfile}. The format of a public
  key should be \texttt{n}, then the
  \texttt{username}, each of which should have been written with a
  trailing newline. The value \texttt{n}
  should have been written as a \emph{hexstring}. See the GMP functions
  for
  \href{https://gmplib.org/manual/Formatted-Input-Functions}{formatted
  input} for help with reading hexstrings.
\end{funcdoc}

% //
% // Generates the components for a new private RSA key.
% // Requires an accompanying RSA public key to complete the pair.
% // All mpz\_t arguments are expected to be initialized.
% //
% // d: will store the RSA private key.
% // e: the precomputed public exponent.
% // p: the first large prime from the public key generation.
% // p: the second large prime from the public key generation.
% //
\begin{funcdoc}{void ss\_make\_priv(mpz\_t d, mpz\_t pq, mpz\_t p, mpz\_t q)}
  Creates a new SS private key \texttt{d} given primes \texttt{p} and
  \texttt{q} and the public key \texttt{n}. To compute \texttt{d}, simply
  compute the inverse of \texttt{n} modulo $\lambda(\texttt{pq}) =
  \lcm(\texttt{p} - 1, \texttt{q} - 1)$.
\end{funcdoc}

\begin{funcdoc}{void ss\_write\_priv(mpz\_t pq, mpz\_t d, FILE
*pvfile)}
  Writes a private SS key to \texttt{pvfile}. The format of a private key
  should be \texttt{pq} then \texttt{d}, both of which are written with a
  trailing newline. Both these values should be written as hexstrings.
\end{funcdoc}

\begin{funcdoc}{void SS\_read\_priv(mpz\_t pq, mpz\_t d, FILE
*pvfile)}
  Reads a private SS key from \texttt{pvfile}. The format of a private
  key should be \texttt{pq} then \texttt{d}, both of which should have
  been written with a trailing newline. Both these values should have
  been written as \emph{hexstrings}.
\end{funcdoc}

\begin{funcdoc}{void ss\_encrypt(mpz\_t c, mpz\_t m, mpz\_t n)}
  Performs SS encryption, computing the ciphertext \texttt{c} by encrypting
  message \texttt{m} using the public key \texttt{n}.
  Remember, encryption with SS is defined as $E(m) = c =
        m^n \pmod{n}$.
\end{funcdoc}

\begin{funcdoc}{void ss\_encrypt\_file(FILE *infile, FILE *outfile,
mpz\_t n)}
  Encrypts the contents of \texttt{infile}, writing the encrypted
  contents to \texttt{outfile}. The data in \texttt{infile} should be in
  encrypted in \emph{blocks}. Why not encrypt the entire file? Because
  of \texttt{n}. We are working modulo \texttt{n}, which means that the
  value of the block of data we are encrypting must be strictly less
  than \texttt{n}. We have two additional restrictions on the values of
  the blocks we encrypt:
  \begin{enumerate}
    \item The value of a block cannot be 0: $E(0) \equiv 0 \equiv 0^n \pmod{n}$.
    \item The value of a block cannot be 1. $E(1) \equiv 1 \equiv 1^n \pmod{n}$.
  \end{enumerate}
  A solution to these additional restrictions is to simply
  \emph{prepend} a single byte to the front of the block we want to
  encrypt. The value of the prepended byte will be \texttt{0xFF}. This
  solution is not unlike the padding schemes such as PKCS and OAEP used
  in modern constructions of RSA\@. To encrypt a file, follow these
  steps:
  \begin{enumerate}
    \item Calculate the block size $k$. This should be
      $k = \lfloor(\log_2({\sqrt{\texttt{n}}}) - 1) / 8\rfloor$.
    \item Dynamically allocate an array that can hold $k$ bytes. This
      array should be of type \texttt{(uint8\_t *)} and will serve as
      the block.
    \item Set the zeroth byte of the block to \texttt{0xFF}. This
      effectively prepends the workaround byte that we need.
    \item While there are still unprocessed bytes in \texttt{infile}:
      \begin{enumerate}
        \item Read at most $k - 1$ bytes in from \texttt{infile}, and
          let $j$ be the number of bytes actually read. Place the read
          bytes into the allocated block starting from index 1 so as to
          not overwrite the \texttt{0xFF}.
        \item Using \texttt{mpz\_import()}, convert the read bytes,
          including the prepended \texttt{0xFF} into an \texttt{mpz\_t
          m}. You will want to set the \texttt{order} parameter of
          \texttt{mpz\_import()} to \texttt{1} for most significant word
          first, \texttt{1} for the \texttt{endian} parameter, and
          \texttt{0} for the \texttt{nails} parameter.
        \item Encrypt \texttt{m} with \texttt{ss\_encrypt()}, then
          write the encrypted number to \texttt{outfile} as a
          \emph{hexstring} followed by a trailing newline.
      \end{enumerate}
  \end{enumerate}
\end{funcdoc}

\begin{funcdoc}{void ss\_decrypt(mpz\_t m, mpz\_t c, mpz\_t d,
mpz\_t pq)}
  Performs SS decryption, computing message \texttt{m} by decrypting
  ciphertext \texttt{c} using private key \texttt{d} and public modulus
  \texttt{n}. Remember, decryption with SS is defined as $D(c) = m =
        c^d \pmod{pq}$.
\end{funcdoc}

\begin{funcdoc}{void ss\_decrypt\_file(FILE *infile, FILE *outfile,
mpz\_t pq, mpz\_t d)}
  Decrypts the contents of \texttt{infile}, writing the decrypted
  contents to \texttt{outfile}. The data in \texttt{infile} should be
  decrypted in \emph{blocks} to mirror how \texttt{ss\_encrypt\_file()}
  encrypts in blocks. To decrypt a file, follow these steps:
  \begin{enumerate}
    \item Dynamically allocate an array that can hold
        $k = \lfloor(\log_2({\texttt{pq}}) - 1) / 8\rfloor$ bytes. This
      array should be of type \texttt{(uint8\_t *)} and will serve as
      the block.
          \begin{itemize}
              \item We need to ensure that our buffer is able to hold at
                  least the number of bits that were used during the
                  enryption process.  In this context we don't know the
                  value of \texttt{n}, but we can overestimate the number
                  of bits in $\sqrt{n}$ using \texttt{pq}.\\
                  \[\log_2(\sqrt{n}) = \log_2(\sqrt{p^2\times q})
                  = \log_2(p\times \sqrt{q}) < \log_2(pq)\]
          \end{itemize}
    \item Iterating over the lines in \texttt{infile}:
      \begin{enumerate}
        \item Scan in a hexstring, saving the hexstring as a
          \texttt{mpz\_t c}. Remember, each block is written as a
          hexstring with a trailing newline when encrypting a file.
        \item First decrypt \texttt{c} back into its original value
          \texttt{m}.  Then using \texttt{mpz\_export()}, convert \texttt{m} back into
          bytes, storing them in the allocated block. Let $j$ be the
          number of bytes actually converted. You will want to set the
          \texttt{order} parameter of \texttt{mpz\_export()} to
          \texttt{1} for most significant word first, \texttt{1} for the
          \texttt{endian} parameter, and \texttt{0} for the
          \texttt{nails} parameter.
        \item Write out $j - 1$ bytes starting from index 1 of the block
          to \texttt{outfile}. This is because index 0 must be prepended
          \texttt{0xFF}. Do not output the \texttt{0xFF}.
      \end{enumerate}
  \end{enumerate}

\end{funcdoc}

% No signing right?
% \begin{funcdoc}{void rsa\_sign(mpz\_t s, mpz\_t m, mpz\_t d, mpz\_t
% n)}
%   Performs RSA signing, producing signature \texttt{s} by signing
%   message \texttt{m} using private key \texttt{d} and public modulus
%         \texttt{n}. Signing with RSA is defined as $S(m) = s = m^d \pmod{n}$.
% \end{funcdoc}

% \begin{funcdoc}{bool rsa\_verify(mpz\_t m, mpz\_t s, mpz\_t e,
% mpz\_t n)}
%   Performs RSA verification, returning \texttt{true} if signature
%   \texttt{s} is verified and \texttt{false} otherwise. Verification is
%         the inverse of signing. Let $t = V(s) = s^e \pmod{n}$. The signature is
%   verified if and only if $t$ is the same as the expected message
%   \texttt{m}.
% \end{funcdoc}
