\section{Introduction}

\epigraphwidth=0.65\textwidth
\epigraph{\emph{I wonder why it is that when I plan a route too
    carefully, it goes to pieces, whereas if I blunder along in blissful
    ignorance aimed in a fancied direction I get through with no
    trouble.}}{---John Steinbeck, \emph{Travels with Charley: In Search
    of America}}

\noindent
Denver Long decided to augment his income during his retirement
years by selling the prestigious products produced by the Shinola
Corporation.  He enjoys driving his Cadillac, so it's the life of
a traveling salesman for him.  He loads up his little dog Satan---whom
he calls \emph{Baby}---and heads out on his new career.

But his first trip does not go so well. Heading to his son's house
after visiting his new grandson, he accidentally takes a wrong turn
near Chula Vista and winds up in Mexico.  Despite many pleasant
visits to Tijuana in years past, visiting his old friend Se\~nor Vasquez
on his ranchero, shooting their rifles at an old El Dorado that he has traded to Se\~nor
Vasquez decades before,
it's no longer the familiar Mexico of 1974.  Having
no passport, and speaking very little Spanish, he turns around in
frustration.  The Border Patrol won't let him back into the United
States for many hours, until he finally wears them down through his
power of persuasion.

Since the profit of his new enterprise depends
on the cost and duration of travel, and losing a day in Tijuana
cost him a potential sale in Barstow, he asks his eldest son to
have his class create a computer program that will provide an optimal
route to all the cities along his way and then return him to his
home in scenic Clearlake.
