\section{Deliverables}
\epigraphwidth=0.6\textwidth
\epigraph{\emph{Self pity is easily the most destructive of the non-pharmaceutical narcotics; it is addictive, gives momentary pleasure and separates the victim from reality.}}{---John Gardner}

\noindent You will need to turn in the following source code and header files:

\begin{enumerate}
  \item \texttt{encode.c}: This file will contain your implementation of the
    Huffman encoder.

  \item \texttt{decode.c}: This file will contain your implementation of the
    Huffman decoder.

  \item \texttt{defines.h}: This file will contain the macro definitions
    used throughout the assignment. \textcolor{red}{You \emph{may not}
    modify this file.}

  \item \texttt{header.h}: This will will contain the \texttt{struct}
    definition for a file header. \textcolor{red}{You \emph{may not}
    modify this file.}

  \item \texttt{node.h}: This file will contain the node ADT interface.
    This file will be provided. \textcolor{red}{You \emph{may not}
    modify this file.}

  \item \texttt{node.c}: This file will contain your implementation of
    the node ADT.

  \item \texttt{pq.h}: This file will contain the priority queue ADT
    interface. This file will be provided. \textcolor{red}{You \emph{may
    not} modify this file.}

  \item \texttt{pq.c}: This file will contain your implementation of the
    priority queue ADT. You \emph{must} define your priority queue
    \texttt{struct} in this file.

  \item \texttt{code.h}: This file will contain the code ADT interface.
    This file will be provided. \textcolor{red}{You \emph{may not}
    modify this file.}

  \item \texttt{code.c}: This file will contain your implementation of
    the code ADT.

  \item \texttt{io.h}: This file will contain the I/O module interface.
    This file will be provided. \textcolor{red}{You \emph{may not}
    modify this file.}

  \item \texttt{io.c}: This file will contain your implementation of the
    I/O module.

  \item \texttt{stack.h}: This file will contain the stack ADT
    interface. This file will be provided. \textcolor{red}{You \emph{may
    not} modify this file.}

  \item \texttt{stack.c}: This file will contain your implementation of
    the stack ADT. You \emph{must} define your stack \texttt{struct} in
    this file.

  \item \texttt{huffman.h}: This file will contain the Huffman coding
    module interface. This file will be provided. \textcolor{red}{You
    \emph{may not} modify this file.}

  \item \texttt{huffman.c}: This file will contain your implementation
    of the Huffman coding module interface.
\end{enumerate}


You can have other source and header files, but \emph{do not try to be
overly clever}. You will also need to turn in the following:

\begin{enumerate}

  \item \texttt{Makefile}: This is a file that will allow the grader to
    type \texttt{make} to compile your programs.

    \begin{itemize}
      \item \texttt{CC = clang} must be specified.

      \item \texttt{CFLAGS = -Wall -Wextra -Werror -Wpedantic}
        must be included.

      \item \texttt{make} should build the encoder and the decoder, as
        should \texttt{make all}.

      \item \texttt{make encode} should build \emph{just} the encoder.

      \item \texttt{make decode} should build \emph{just} the decoder.

      \item \texttt{make clean} must remove all files that are compiler
        generated.

      \item \texttt{make format} should format all your source code,
        including the header files.
    \end{itemize}

  \item Your code must pass \texttt{scan-build} \emph{cleanly}. If there
    are any bugs or errors that are false positives, document them and
    explain why they are false positives in your \texttt{README.md}.

  \item \texttt{README.md}: This must be in \emph{Markdown}. This must describe
    how to build and run your program.

  \item \texttt{DESIGN.pdf}: This \emph{must} be a PDF\@. The design document
    should answer the pre-lab questions, describe the purpose of your program,
    and communicate its overall design with enough detail such that a sufficiently
    knowledgeable programmer would be able to replicate your implementation.
    \textcolor{red}{This does not mean copying your entire program in verbatim.}
    You should instead describe how your program works with supporting pseudocode.
    \textcolor{red}{\textbf{C} code is \textbf{not} considered pseudocode.}
\end{enumerate}
